A Distributed and Tensor-based Approach for Multidimensional Dictionary Learning

### 2017-07-12 docs/papers/2016JNCA/submition/response02/jnca-response_to_reviewers.txt

Response to the Handling Editor and Reviewers

Submission No.: JNCA-D-16-00617R1
Title: Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks
Journal of Network and Computer Applications (JNCA)
Research Paper

Dear Mohammed Atiquzzaman, Ph.D., Editor-in-Chief, and reviewers,
The authors would like to thank the Associate Editor-in-Chief and the Reviewers for volunteering their time in reviewing our manuscript and providing us with valuable comments which allowed us to significantly improve the paper. We have carefully revised the manuscript based on the reviewer’s comments and suggestions. All changes in the paper have been marked with a color. The following is our point-by-point responses to the raised comments.

Kind regards,

The authors.

1. EDITOR
1.1. COMMENT: The referees have suggested that certain aspects of the paper should be re-examined, before it can be considered further for publication.
1.1.1. REPLY: We thank the Editor for the time, work, and positive evaluation of our paper.

2. REVIEWER #3
2.1. COMMENT: The reviewer appreciates the efforts that the authors have made to revise this paper. The revised version has improved a lot compared to the previous version. Most of the questions raised from the reviewers have been clarified. The reviewer has three additional comments regarding the revised version.
2.1.1. REPLY: Obrigado pela revisão positiva de maneira geral. Esperamos poder responder os comentários adicionais da maneira esperada.
2.2. COMMENT: The mentioned dataset used for additional evaluation is almost twenty years ago, and it might be able to reflect the current state of arts of modern attacks.
2.2.1. REPLY: Entendemos esta limitação e até inclimos comentários sobre isto durante a revisão da literatura. Mesmo este dataset sendo velho e podendo nao refletir ataques modernos, este dataset ainda vem sendo utilizado por lacunas em datasets novos e públicos, conforme surveys recentes.
2.3 COMMENT: The authors have provided complexity and running time evaluation in the revision. However, given that the EVD complexity in O(M^2*N+N^3), it looks like the proposed approach might be DoS'ed if an attacker attempts to flood using randomized port numbers, e.g., (M=) 64K network ports. This is because performing bandwidth consumption flooding attacks does not need to work with a dedicated (or a monitored) port number.
2.3.1. REPLY: A proposta é a aplicação da técnica em portas selecionadas de acordo com seu interese e seu uso em uma eventual rede interna. Nossos cenários consideraram ataques provinientes de de uma rede interna e também de uma rede externa para uma rede interna. "performing bandwidth consumption flooding attacks" não são algo da nossa análise, mas nossa técnica foi capaz de detectar flood atack em casos que utilizam diveras portas , utilizando análise de correlação, como nos casos avaliados para DARPA 98 nas seções X e Y
2.4 COMMENT: The authors have revised the paper to focus exactly on probe and scanning attacks. This may greatly limited the contribution of the proposed work, as we already have a number of commercial solutions to deal with the problem. 
2.4.1. REPLY: Nosso trabalho foca em flooding, probe and scannnig, buscando desenvolver uma nova tecnica sem dependência de conhecimento prévio, análise de comportamentos conhecidos, dicionários, e etc. Nosso objetivo é utilizar uma téction unsupervisioned e blind

3. REVIEWER #4
3.1. COMMENT: The authors have satisfactorily addressed the queries raised by the reviewers. However, only DARPA 1998 dataset has been considered to conduct the experiments. The authors can also conduct the experiments on other publicly available datasets such as CAIDA (https://www.caida.org/data/) and UNB-ISCX (http://www.unb.ca/research/iscx/dataset/iscx-IDS-dataset.html). Such datasets contain more recent attack vectors and types. Conducting experiments on such datasets and discussion of analysis on such datasets can provide in depth understanding of the proposed work.
3.1.1. REPLY: Obrigado pela construtuva comentário... Estes datasets foram considerados e citados em X e Y, entretanto os ataques monitorados por estes casos não consideram probe attack e flooding em portas tcp/udp, como também são baseados em fluxo enquanto nossa técnica é baseada em pacotes. Por fim, mas nao menos importante, estes datasets não dispõem de uma classificação prévia (label) dos ataques. Como UNB-ISCX

4 REVIEWER #5
4.1. COMMENT: Since only one user (computer) was used, the amount of traffic noise (DHCP) would be very limited.
4.1.1. REPLY: Como forma de adicionar cenários mais abrangentes e realistas e bem conhecidos, adicionamos a análise da nossa tecnica aplicada ao DARPA 98
4.2. COMMENT: There is no description of the systems used, because attacks such as the fraggle attack are easily eliminated by both systems and routers. Plus, the network is very limited with the number of devices used and type of traffic observed.
4.3.1. REPLY: Para o fraggle utilizamos a ferramenta X, uma descrição sobre isto foi adicionada no em Z
4.4. COMMENT: The authors stated """The UDP scan technique sends UDP packets to the destination port, and if it responds with a ICMP port 230 unreachable message, then it indicates that the scanned port is closed. On the other hand, if a message is not received, then the port is considered as open."""* lots of systems today disable this by default, that is why it is unclear which systems were used for testing.
4.4.1. REPLY: Esta descrição se refere a uma forma de portscan avaliada em nosso cenário sintético, mas os cenários de portscan avaliados no dataset darpa incluem casos em que não há resposta para o envio de pacotes
4.5. COMMENT: What if the port scan was not sent in a sequence fashion and sent with time frames higher than the samples selected for evaluation and checking? The information considered in the proposed framework will not help.
4.5.1. REPLY: Casos como este foram avaliados e não foram identificados, conforme descritos em XX
4.6. COMMENT: Why didn't the authors use flow analysis or even compare to such related work. Flow analysis has been used a lot for such detection, especially when you're looking at time and ports.
4.6.1. REPLY: Flow analysis implica em custo adicional para a identificação e montagem de fluxos para uma eventual análise. Entretanto, iremos considerar uma análise baseada em fluxo como trabalhos futuros.
4.7. COMMENT: How can such module detect new novel attacks, when the only experiments tested were SYN(flood), fraggle (rare today), and two port scanning techniques? ex: traffic like DHCP are considered noise, so how would this framework detect an ARP Scan? O framework proposto foca em flood e probe attacks na camada de transporte, entretanto a nossa técnica pode highlight ataques inesperados que tenham comportamentos semelhantes ao que selecionados, sendo então possível selecionar executar análises mais detalhadas nestes casos específicos. Dado que ARP scan foca em uma camada abaixo da camada de transporte, este não seria detectado por nossa proposta
4.7.1. REPLY: 
4.8. COMMENT: The Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack shows that port 80, 443, and 53 were not an attack, even though all of these traffic are currently used for exfiltration or covert communication. So, how will the proposed framework detect such attacks? Detecting such attacks could be considered novel if they could achieve such detection.
4.8.1. REPLY: No cenário considerado o tráfego nas portas 80, 443, and 53 foi modelado como legítimo, representando tráfego de usuários aos serviços HTTP, HTTPS e DNS para demonstrar a capacidade de detecção de ataques mixed into legitimate traffic. Nosso framework não é capaz de detectar exfiltration, covert communication ou exploiting, como demonstrado ao analisarmos nossa solução para o dataset DARPA. Podemos considerar iste ponto como um foco para nosso trabalho de doutorado, com uma análise de correlação para identificar casos de exfiltration or covert communication

### 2017-07-12 docs/papers/2017EUSIPCO/2017eusipco.tex
returning dictionary and sparse coding. implementation of method evaluation saving into files

\usepackage{graphicx}
\usepackage{array}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage[mathletters]{ucs}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{morefloats}
\usepackage{lipsum}
\usepackage{amsthm,amsmath}

\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms

@@

\title{A Distributed Tensor-based Approach for Multidimensional Dictionary Learning}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Thiago P de B Vieira}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Andr ́e L. F. de Almeida}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{Florian Roemer\\ and João Paulo Lustosa}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

@@

\section{Introduction}
Dictionary Learning is a signal processing technique useful for sparse representation of signals throught the estimation of basis vectors and learning sparse representations of training data. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and in solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing.

In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly bahavior of users in social networks or when evaluating correlation between audio and video. Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in data analytics problems. Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}. However, tensor-based algorithms face challenges to achieve reasonable performance to handle large-scale tensor factorizations, demanding efforts in order to explore distributed techniques for tensor-based big data analytics.

	Distributed tensor-based...

	Distributed tensor based dictionary learning...

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_1000_100.eps}
     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

@@

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix and $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise.

where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$.

We adopt the signal-to-noise ratio (snr) of 20 and the sparseness factor of 5 in order to generate the separable dictionary $\boldsymbol{A}$ according to $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$ of a gaussian distributed data $\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}$.

\begin{equation}\label{eq:eq04}
\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
\end{equation}

>>>>

%% packages
% graph
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{array}
\usepackage{rotating}
\usepackage{tabularx}
% math
\usepackage{amsfonts}
\usepackage{mathtools}
% algorithm
\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms

@@

\title{A Distributed and Tensor-based Approach for Multidimensional Dictionary Learning}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Thiago P de B Vieira}
\IEEEauthorblockA{
	Department of Electrical Engineering\\
	University of Brasilia\\
	Bras\'ilia-DF, Brazil\\
	Emai	l: tpbvieira@gmail.com
}

\and

\IEEEauthorblockN{Andr\'e L. F. de Almeida}
\IEEEauthorblockA{
	Department of Teleinformatics Engineering\\
	Federal University of Cear\'a\\
	Fortaleza, Brazil\\
	Emai	l: andre.almeida@ieee.org
}

\and

\IEEEauthorblockN{Florian Roemer\\ and Jo\~ao Paulo C. L. da Costa}
\IEEEauthorblockA{
	Institute for Information Technology\\
	Ilmenau University of Technology\\
	Ilmenau, Germany
}

}

@@

\section{Introduction}

Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.

In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks. Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}. Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.

Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems. Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.

Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results. However, the performance of tensor-based algorithms outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the detection rating of traditional algorithms decreases over the dataset increasing.

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_1000_100.eps}
     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
     \label{fig:fig1}
\end{figure}

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
     \label{fig:fig2}
\end{figure}

Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data. Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries. It is possible to observe better detection rating for tensor-based algorithms. However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.

We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets. We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition. We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.

Consider a generic sparse recovery problem of the following form

where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{•}hbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.

Consider a sparse recovery problem for a separable 2-D manifold, i.e., a manifold that can be written as the product of two 1-D manifolds. If we choose a separable 2-D sampling grid, we can write the 2D sparse recovery problem as

where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.

The Kronecker model (\ref{eq:eq02}) can be rewritten in an equivalent tensor form. Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:eq02}) into a "Tucker-2" decomposition, as

We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained. In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons. We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.

The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.

...

...

@@
